---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
library(janitor)
```

```{r}
qb_revenue_breakdown
rev <- clean_names(qb_revenue_breakdown)
```

```{r}
qb_monthly_sales
```
```{r}
qb_competitors
```

```{r}
qb_device_data
```
Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.
```{r}
qb_revenue_breakdown %>% 
  ggplot() +
  aes(x = Year, y = Revenue, fill = Product) +
  geom_col()
```
Make a line chart showing monthly sales of the “flagship product” - the ‘KwikBit Unit 1’ - over the last year. This data is inside qb_monthly_sales, but you will need to use filter() or subset() (base R).
```{r}
qb_monthly_sales %>% 
  filter(Cashflow == "Kwikbit Unit 1 Sales") %>% 
  ggplot() +
  aes(x = Date, y = Amount) +
  geom_line()
```

Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales
```{r}
qb_monthly_sales %>% 
  filter(Cashflow %in% c("Total Revenue", "Cost")) %>% 
ggplot() +
  aes(x = Date, y = Amount, group = Cashflow, colour = Cashflow) +
  geom_line()
```

Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.

```{r}
qb_competitors %>% 
  ggplot() +
  aes( x = Year, y = Revenue, group = Company, fill = Company) +
  geom_area()
```

Now show the sales from the four competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
qb_competitors %>% 
  ggplot() +
  aes( x = Year, y = Revenue, group = Company, colour = Company) +
  geom_line() +
  geom_point()
```

Now the company wants to compare the number of steps that their device counts vs. their competitors. Make a line graph of the number of steps throughout time, and use faceting to compare between companies and people. The data for this is in qb_device_data.

```{r}
qb_device_data %>% 
  ggplot() +
  aes(x = hours, y = counts, group = companies) +
  geom_line() +
  facet_grid(~ companies)
```

