---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```
Question 1.

Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.
```{r}
summary(backpack)

backpack
```

Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.
```{r}
backpack %>% 
  ggplot() +
  aes(x = backpack_weight) +
  geom_bar()
```
```{r}

backpack %>% 
  ggplot() +
  aes(x = backpack_weight) +
  geom_histogram(binwidth=.5, colour="black", fill="white") +
    geom_vline(aes(xintercept=mean(backpack_weight, na.rm=TRUE)),   # Ignore NA values for mean
               color="red", linetype="dashed", size=1)
```
Question 3.

Are there differences in backpack weight between sex?


```{r}
backpack %>% 
  ggplot() +
  aes(x = backpack_weight, fill = sex) +
  geom_bar() 


```


Question 4.

What is the relationship between body weight and backpack weight?

```{r}
backpack %>% 
  ggplot() +
  aes(x = body_weight, y = backpack_weight, colour = sex) +
  geom_point()
```

Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

Seems that there is a correlation between body_weight and backpack_weight ie heavier body weight the larger the backpack weight. THis aligns to men and women.

```{r}
backpack %>% 
  ggplot() +
  aes(x = body_weight, y = backpack_weight, colour = sex) +
  geom_point()
```


Question 6.

What proportion of students report back pain?

```{r}
# backpack %>% 
#   group_by(back_problems) %>% 
#   summarise(count = n()) %>% 
  ggplot(backpack) +
  aes(x = back_problems) +
  geom_bar() +
    labs(
      x = "back problems",
      y = "count",
      title = "Proportion with and without back problems") +
  scale_x_discrete() +
  theme_minimal()

```
Question 7.

Is this proportion different between men and women?


```{r}
ggplot(backpack) +
  aes(x = back_problems, fill = sex) +
  geom_bar() +
    labs(
      x = "back problems",
      y = "count",
      title = "Proportion with and without back problems") +
  scale_x_discrete() +
  theme_minimal()
```

Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}

```

