---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
library(gtools)

```


#total is 7
```{r}

rolldie(2) %>% 
  dplyr::filter(X1 + X2 == 7)
```

```{r}
rolldie(2) %>% 
 dplyr::filter(X1 + X2 >= 10)

```

#%% is one devided by each that has no remainder - modula operator - use to give the raminder 

#in this the remainder is 0 then it is an even number
```{r}
rolldie(2) %>% 
  dplyr::filter(X1 %% X2 == 0 )
```

```{r}
1/12
```

```{r}
merge(rolldie(1),tosscoin(1)) %>% 
  dplyr::filter(X1 %% 2 == 0 & toss1 == "T")
```

```{r}

S_rolldie_tosscoin <- merge(rolldie(1), tosscoin(1))
die_is_even_coin_is_tails <- S_rolldie_tosscoin %>% 
  dplyr::instafilter(X1 %% 2 == 0 & toss1 == "T")
nrow(die_is_even_coin_is_tails)/nrow(S_rolldie_tosscoin)
```

```{r}
permutations(
  n = 4,
  r = 4,
  v = c(2, 3, 4, 5),
  repeats.allowed = FALSE
  )


```


#any 4 digit numbers are there
```{r}
locker_numbers <- permutations(
  n = 10,
  r = 4,
  v = 0:9,
  repeats.allowed = TRUE
)
```

```{r}
choose(55,6)

dim(lottery)
```

```{r}
employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Gabra", "Hua", "Inigo", "Hugo")

seniors <- permutations(
  n = 10,
  r = 3,
  v = employees,
  repeats.allowed = FALSE
)

dim(seniors)
```

```{r}
#as_tibble better as displaying - then set_names changes all the headers in columns 

as_tibble(seniors) %>% 
  set_names(c("CEO", "CTO", "COO")) %>% 
  dplyr::filter(CEO == "Alice" & CTO == "Bob")
```

Task - 10 mins A licence plate begins with three letters from a possible set of A, B, C, D, E. How many permutations are possible if repetition of letters is:
not allowed [Hint - how many letters are possible for first position, how many for second if one is ‘used up’, how many for third if two are ‘used up’?]
allowed [Hint - all that changes from last answer is that letters don’t get ‘used up’]
Try this either using pen and paper maths or R, your choice!

```{r}
letters <- c("A", "B", "C", "D", "E")

permutations(
  n = 5,
  r = 3,
  v = letters,
  repeats.allowed = FALSE
)

permutations(
  n = 5,
  r = 3,
  v = letters,
  repeats.allowed = TRUE
)

```

```{r}

all_cards <-
cards() %>% 
  mutate(card = str_c(rank, suit)) %>% 
  pull (card)


```


#permutations and combinations - combinations are use when you dont matter what order they appear in.
```{r}

five_card_hands <- 
combinations(
  n = 52,
  r = 5,
  v = all_cards,
  repeats.allowed = FALSE
)
```

```{r}
options(scipen = 999)
dim(five_card_hands)
```

```{r}
4/2598960
```


#a number of ways to pick 5 cards from 52 - 
```{r}

choose(52,5)
```

```{r}
lottery_tickets <- 
combinations(
  n = 50,
  r = 6,
  v = 1:50,
  repeats.allowed = FALSE
)
```

```{r}

dim(lottery_tickets)
```

I have five flavours of ice cream to choose from:
# SLACK THIS OUT
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio") 
and a bowl that can hold three scoops. How many different variations in bowls of ice cream can I have assuming:
Each scoop must be a different flavour
I can repeat flavours if I wish?


```{r}

#number of combinations
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")

icecream <- combinations(
  n = 5,
  r = 3,
  v = flavours,
  repeats.allowed = FALSE
)

dim(icecream)

```


# use choose for finding the number of combinations possible 
```{r}

choose(5, 3)
```

```{r}
tosscoins(4) %>% 
  muatate(num_heads)
```


#sometimes easier to calculate the opposit and take it away from 1 
# known as compliments

#P(Ac)= 1-P(A)

#intersections - outcomes common to both event spaces 


Let’s check the addition law for ourselves. Calculate probabilities for each event below:
first_die_even
second_die_5
first_die_even AND second_die_5
Then use these probabilities in the expression from above
prob(A∪B)=prob(A)+prob(B)−prob(A∩B)
and check you get the same probability for prob(first_die_even OR second_die_5)
prob(first_die_even OR second_die_5) as we obtained from the last block of code.

```{r}
rolldie(2)


```

